Untwisting the Tetrahelix (v0.4)

### 2017-02-09 doc/TetrahelixUntwist.tex
lambda continuum working

\title{Untwisting the Tetrahelix (v0.4)}
\author{Robert L. Read
  \thanks{read.robert@gmail.com}
}
\affil{Founder, Public Invention, an educational non-profit.}

@@

Just as $H_{general}$ constructs the BC helix (with careful and non-obvious choices of parameters) which is an important
special case due to its regularity, it constructs an additional special (degenerate) case when the rail angle $\rho = 0$
and $h = 1$ (the edgelength), which we call the \emph{equitetrabeam}.

Note: place here a computation of the length classes, and a means of computing from $H_{general}$.
We should have a close-form expression for the short, mid-length, and long classes.

By theorem \ref{generalformulaoptimal}, the equitetrabeam is optimal in terms of its lengths.

\begin{theorem}
  The equitetrabeam is also optimal under the metric of the total length of all members 
  for an equilateral tetrabream isomorphic to the Boerdijk--Coxeter helix.
\end{theorem}

\begin{proof}
  Summing together all lengths:
  \begin{align*}
  sum &= orange_e + orange_o + purple_e + purple_e + purple_o + green_e + green_o \\    
  &= \sqrt{1 + Y^2} + \sqrt{1 + (1-Y)^2} + \sqrt{1 + B^2} + \sqrt{1+ (1-B)^2} + \\
  & \sqrt{1 + (B - Y)^2} +  \sqrt{1 + ((1+Y) - B)^2} \\
  \end{align*}

  Wolfram Alpha, evaluating this numerically, gives the same minima:

  \begin{tabular}{c}    
$  min\{\sqrt(1 + Y^2) + \sqrt{1 + (1 - Y)^2} + \sqrt{1 + B^2} + \sqrt{1 + (1 - B)^2} + $\\
    $  \sqrt{1 + (B - Y)^2} + \sqrt{1 + ((1 + Y) - B)^2}\} $ \\
    $ \approx 6.76783 \text{at } (B, Y) \approx (0.666667, 0.333333) $
  \end{tabular}

@@

>>>>

\title{Untwisting the Tetrahelix (v0.4)}
\author{Robert L. Read \texttt{read.robert@gmail.com} \and
  Robert Gatliff \texttt{robert@toubat.org}
}

@@

Just as $H_{general}$ constructs the BC helix (with careful and non-obvious choices of parameters) which is an important
special case due to its regularity, it constructs an additional special (degenerate) case when the rail angle $\rho = 0$
and $d = 1$ (the edgelength), where the cross sectional area is an equilateral triangle of unchanging orientation.
We call this the \emph{equitetrabeam}.

However, choosing $d = 1$ and $\rho = 0$ does fix the radius. We would like to choose the radius so that the
minimax difference between the lengths is as small as possible.

When $\rho = 0$, our formulae for distances are particular simple. Let $a$ be the length of the equilateral triangle
forming the beam cross section. Then by geometry $r = a \sqrt(3)$.
\begin{align*}
  \text{rail} &=  1 \\
  \text{one-hop} &= \sqrt{\frac{1}{3}^2 + (\frac{a}{\sqrt{3}})^2 + (a/2)^2}\\
\text{one-hop}  &= \sqrt{\frac{1}{9} + \frac{7 a^2}{12}} \\
    \text{two-hops} &= \sqrt{\frac{2}{3}^2 + (\frac{a}{\sqrt{3}})^2 + (a/2)^2}  \\
\text{two-hops}    &= \sqrt{\frac{4}{9} + \frac{7 a^2}{12}}
\end{align*}

Note that two-hops increases more and more slowly as $a$ increases. Therefore the minimum maximum will
be obtained with the value of $a$ that when $one-hop = 1$. This can be observed from the
plot of two-hops - one-hops as a function of $a$, or by observing that the derivate of two-hops - one-hop
is negative for $a > 0$. (That is, the value of one-hop increases faster than two-hops.)

We seek the value of $a$ that minimizes the maximum difference between these three lengths. We find
by numerical experimentation that this difference is minimized that the one-hop distance is equal to $1$,
which also has the pleasing property of giving us only two lengths. Then
\begin{align*}
   1  &=  \sqrt{\frac{1}{9} + \frac{7a^2}{12}} \\
     a  &=  \sqrt{\frac{8 \cdot 12}{9 \cdot 7}} \\
          a  &=  \sqrt{\frac{32}{21}}
\end{align*}

Therefore $ r = \sqrt{\frac{32}{63}} \approx 0.71267 $, producing length difference in the two-hop rail of
$\frac{2}{\sqrt{3}}$
and $1$ of $\approx 15.47\% $.

Note: Another interesting solution occurs at $a = \sqrt(5)/2$, which produces three length
classes of $11/12, 12/12, 13/12$.

\begin{conjecture}
  The equitetrabeam is also optimal under the metric of the total length of all members 
  for an equilateral tetrabream isomorphic to the Boerdijk--Coxeter helix.
\end{conjecture}

\begin{proof}

@@

We can then create a continuum of radius, length, rho to place into our general equation
to create continuum of helices between the BC and equitetrabeam. This continuum is 
minimax-length optimal at its endpoints, and may be optimal throughout.
